Seniority and hierarchy configuration interaction for radicals and excited states

### 2023-06-01 Manuscript/manuscript.tex
saving work

@@

hCI1/HF
hCI1.5/HF
hCI2/HF
$\Delta$hCI1
$\Delta$hCI1.5
$\Delta$hCI2

Higher-order levels for a few systems
hCI2.5/HF
hCI3/HF
$\Delta$hCI2.5
$\Delta$hCI3
$\Delta$hCI3.5

>>>>

From here on, method names carrying the $\Delta$ symbol denote a state-specific approach,
whereas those without the $\Delta$ mean that ground-state HF orbitals were employed.

@@

In Tab.~\ref{tab:1} we present the statistical errors of the different methods, organized by their computational scaling.
We found that $\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat more favourable MSE and RMSE of the latter.
$\Delta$hCI2 still tends to underestimate the excitation energies, though to a much lesser extent than observed for the lower-order hCI methods.

\begin{table}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
for the Set of 60 Excitation Energies Listed in the {\SupInf}.
}
\label{tab:1}
\begin{ruledtabular}
\begin{tabular}{lddd}
method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
\hline
$\Delta$CSF       & -0.54 & 0.70 & 0.86 \\
CIS               & +0.13 & 0.59 & 0.77 \\
\hline
hCI1              & +1.11 & 1.14 & 1.39 \\
$\Delta$hCI1      & -0.90 & 1.17 & 1.40 \\
\hline
hCI1.5            & +1.88 & 1.88 & 2.20 \\
$\Delta$hCI1.5    & -1.99 & 2.12 & 2.58 \\
\hline
$\Delta$CISD      & -0.08 & 0.17 & 0.22 \\
$\Delta$hCI2      & -0.15 & 0.21 & 0.30 \\
\hline
sCI2              &  0.96 & 1.06 & 1.28 \\
$\Delta$sCI2      &  0.39 & 0.60 & 0.73 \\
%\hline
%CIS(D)            & -0.03 & 0.35 & 0.40 \\
%CC2               & -0.05 & 0.32 & 0.37 \\
%CCSD              & +0.01 & 0.08 & 0.09 \\
%CC3               & +0.01 & 0.03 & 0.06 \\
%CCSDT             & -0.01 & 0.02 & 0.02 \\
\end{tabular}
\end{ruledtabular}
\end{table}

Despite the rather poor performance of lower-order hCI methods for excitation energies,
we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
For that, we performed calculations for a subset of 14 excited states of small molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
with corresponding statistical errors shown in Tab.~\ref{tab:2}.
Even though this is a small set, it should be enough to obtain the main trends on the statistics of the different methods.

\begin{table}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
for the Set of 14 Excitation Energies Listed in the {\SupInf}.
}
\label{tab:2}
\begin{ruledtabular}
\begin{tabular}{lddd}
method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
\hline
%CIS               & +0.16 & 0.82 & 0.96  \\
%$\Delta$CSF       & -0.84 & 0.91 & 1.03  \\
%\hline
CISD              & +4.09 & 4.09 & 4.18  \\
hCI2              & +2.89 & 3.00 & 3.28  \\
$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
$\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
\hline
hCI2.5            & +1.93 & 1.93 & 2.05  \\
$\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
\hline
CISDT             & +0.09 & 0.18 & 0.20  \\
hCI3              & +0.18 & 0.18 & 0.19  \\
$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
$\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
\hline
$\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
\hline
CISDTQ            & +0.14 & 0.14 & 0.16  \\
$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
\end{tabular}
\end{ruledtabular}
\end{table}

We found that the errors of both ground-state based hCI and state-specific $\Delta$hCI to systematically decrease when going toward higher orders of $h$, which would be expected.
%Unfortunately, the pace of decrease is too slow to render these approaches very useful.
%In line with what has been found for excitation-based CI \cite{Kossoski_2023},

The case $h=2$ was discussed above for the larger set of excitation energies.
Here we just notice that hCI2 is found to be somewhat more accurate than CISD, even though the errors are still too large.
However, this improvement does not reflect on the state specific route (compare $\Delta$hCI2 and $\Delta$CISD in Tabs.~\ref{tab:1} and ~\ref{tab:2}).
For $h=2.5$, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
it is actually less accurate then $\Delta$hCI2, and therefore not useful.
Moving to $h=3$, we

hCI2.5
hCI3
$\Delta$hCI2.5
$\Delta$hCI3
$\Delta$hCI3.5
